---
title: "PA1_template1"
author: "Jose Carlos"
date: "August 2, 2020"
output: 
  
  html_document: default
  md_document: default
---

## What is mean total number of steps taken per day?

```{r reading data, include=FALSE}
data <- read.csv("activity.csv")
library(ggplot2)
library(dplyr)

data$date <- as.Date(data$date, "%Y-%m-%d")
weekday <- weekdays(data$date)
activity <- cbind(data,weekday)
```

# Histogram

```{r nmean}

newdata <- summarize(group_by(data, date), sum(steps))
names(newdata) <- c("date", "steps")
newdata <- na.omit(newdata)
hist(newdata$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", 
     col = "darkblue", ylim = c(0,20), breaks = seq(0,25000, by=2500))
```

# Mean and median

```{r mean and median}
mean(newdata$steps)

median(newdata$steps)
```

## What is the average daily activity pattern?

# Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days

```{r intervals}

intervaldata <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE) 
names(intervaldata) <- c("interval", "mean")
plot(intervaldata$interval, intervaldata$mean, type = "l", col="darkblue", lwd = 2, 
     xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")

```

# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r max interval}

intervaldata[which.max(intervaldata$mean), ]$interval

```

## Imputing missing values

# Calculate and report the total number of missing values in the dataset

```{r NAs}

sum(is.na(data))

```

# Devise a strategy for filling in all of the missing values in the dataset.

```{r}

imputed_steps <- intervaldata$mean[match(data$interval, intervaldata$interval)]

```

# Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}

imputed_steps <- intervaldata$mean[match(data$interval, intervaldata$interval)]

```

# Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}

activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), 
                                                       yes = imputed_steps, no = activity$steps))
total_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_imputed) <- c("date", "daily_steps")

```

# Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}

hist(total_imputed$daily_steps, col = "blue", xlab = "Total steps per day", ylim = c(0,30), 
     main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))

```

# mean and median 

```{r}

mean(total_imputed$daily_steps)

median(total_imputed$daily_steps)

```


## Are there differences in activity patterns between weekdays and weekends?

# Create a new factor variable in the dataset with two levels: “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r weekdays vs weedend}

data$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
data$datetype <- sapply(data$date, function(x) {
        if (weekdays(x) == "Saturday" | weekdays(x) =="Sunday") 
        {y <- "Weekend"} else 
        {y <- "Weekday"}
        y
})

```

# Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.

```{r weekdays vs weedend plot}

activitydate <- aggregate(steps~interval + datetype, data, mean, na.rm = TRUE)
plot <- ggplot(activitydate, aes(x = interval , y = steps, color = datetype)) +
        geom_line() +
        labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
        facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)

```


